---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  tidy = FALSE,
  error = FALSE,
  fig.width = 8,
  fig.height = 8)
```

# presser

> Your own web server for happy HTTP testing

<!-- badges: start -->
[![R build status](https://github.com/gaborcsardi/pressr/workflows/R-CMD-check/badge.svg)](https://github.com/gaborcsardi/pressr/actions)
[![](http://www.r-pkg.org/badges/version/presser)](http://www.r-pkg.org/pkg/presser)
[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/presser)](http://www.r-pkg.org/pkg/presser)
<!-- badges: end -->

Lightweight web apps for testing. Built using the
[https://github.com/civetweb/civetweb](civetweb) embedded web server.

## Features

* Complete web app framework, define handlers for HTTP request in R.
* Flexible path matching, with parameters and regular expressions.
* Built in templating system using glue or bring your own template engine.
* Middleware to parse JSON, multipart and URL encoded request bodies.
* A web app is just an R object. It can be saved to disk, copied to another
  R process, etc.
* A web app is extensible, by adding new routes and middleware to it.
* Helper functions for sending JSON, files from disk, etc.
* Comes with an app similar to the https://httpbin.org API, so often
  you don't need to write your own web app.
* The web server runs in the R process, so it has no problems with local
  firewalls.
* Write your own app for your custom test cases.
* Run one web app per test suite, per test file or per test case.
* Multi-threaded web server supports concurrent HTTP requests.
* Limit download speed to simulate low bandwidth.

## Optional dependencies

* The jsonlite package is needed for the `mw_json()` middleware, the
`response$send_json()` method and the `httpbin_app()` app.
* The glue package is needed for the `tmpl_glue()` template engine.
* The callr package is needed for `new_app_process()` to work.

## Installation

Once on CRAN, install the package as usual:

```r
install.packages("presser")
```

## Usage

Start a web app at the beginning of your tests or test file, and
stop it after. Here is an example with the testthat package. Suppose you
want to test that your `get_hello()` function can query an API:

```{r include = FALSE}
library(testthat)
```

```r
web <- setup({
  app <- presser::new_app()
  app$get("/hello/:user", function(req, res) {
    res$send(paste0("Hello ", req$params$user, "!"))
  })
  presser::new_app_process(app)
})
teardown(web$stop())

test_that("can use hello API", {
  url <- web$url("/hello/Gabor")
  expect_equal(get_hello(url), "Hello Gabor!")
})
```

When testing HTTP clients you can often use the built in `httpbin_app()`:

```{r}
httpbin <- setup(presser::new_app_process(presser::httpbin_app()))
teardown(httpbin$stop())

test_that("HTTP errors are caught", {
  url <- httpbin$url("/status/404")
  resp <- httr::GET(url)
  expect_error(httr::stop_for_status(resp), class = "http_404")
})
```

## Documentation

See https://r-lib.github.io/presser/

## License

MIT © RStudio
