---
title: "How to use presser in your tests"
output: rmarkdown::html_vignette
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## How do I use presser in my package?

First, you need to add presser to the `DESCRIPTION` file of your package.
Use the `Suggests` field, as presser is only needed for testing:
```
...
Suggests:
  presser,
  testthat
...
```

Then you need to decide if you want a single web app for all your test cases.
The alternative is to use different apps for some or all test files.
Occasionally you may want to use a special app for a single test case.
Each app runs in a new subprocess, and it takes typically about 100-400ms to start.

See the sections below on writing tests with a single app or multiple apps.

## How do I use `httpbin_app()` (or another app) with testthat?

```{R child = "partials/_howto-testthat.Rmd"}
```

## How to make sure that my code works with the _real_ API?

Indeed, if you use presser for your test cases, then they never touch
the real web server.
As you might suspect, this is not ideal, especially when you do not
control the server.
The web service might change their API, and your test cases will fail
to warn you.

One practical solution is to write (at least some) flexible tests,
that can run against a local fake webserver, or a real one, and you
have a quick switch to change their behavior.
I have found that environment variables work great for this.

E.g. if the `FAKE_HTTP_TESTS` environment variable is not set, the
tests run with the real web server, otherwise they use a fake one.
Another solution, that works best is the HTTP requests are in the
downstream package code, is to introduce one environment variable for each
API you need to connect to.
These might be set to the real API servers, or to the fake ones.

Once you have some tests that can use both kinds or servers, you can
set up your continuous integration (CI) framework, to run the tests
agains the real server (say) once a day.
This special CI run makes sure that your code works well with the
real API.
You can run all the other tests, locally and in the CI, against the
fake local web server.

See the next question for how presser helps you setting
environment variables that point to your local server.

## How do I make my app connect to presser when the tests are running?

In the typical scenario, you want your package to connect to the test app only when running the tests.
One way to achieve this is to allow specifying the web server URL(s) via environment variables.
E.g. when writing a GitHub API client, your package can check use `GITHUB_URL` environment variable.
When this is not set, the package connects to the proper GitHub API.
When testing, you can point it to your test app.

`new_app_process()` helps you setting up temporary environment variables.
These are active while the process is running, and they are removed or reset in `$stop()`. For example:

```{r, include = FALSE}
# Make sure this is not set in the vignette
Sys.unsetenv("GITHUB_API")
```

```{r}
http <- presser::new_app_process(presser::httpbin_app())
http$local_env(list(GITHUB_API = http$url()))
Sys.getenv("GITHUB_API")
http$stop()
Sys.getenv("GITHUB_API")
```

## How can I write my own app?

```{r child = "partials/_my-app.Rmd"}
```

## Can I have an app for a single testthat test file?

To run a web app for a single test file, start it in `testthat::setup()` and stop it in `testthat::teardown()`.
You can do this at the beginning of the file:

```{r}
web <- setup({
  app <- presser::new_app()
  app$get("/hello/:user", function(req, res) {
    res$send(paste0("Hello ", req$params$user, "!"))
  })
  presser::new_app_process(app)
})
teardown(web$stop())
```

Then in the test cases, use `web$url()` to get the URL to connect to.

```{r}
test_that("can use hello API", {
  url <- web$url("/hello/Gabor")
  expect_equal(httr::content(httr::GET(url)), "Hello Gabor!")
})
```

## Can I use an app for a single testthat test?

Sure.
For this you need to create the app process within the `testthat::test_that()` test case and use `on.exit()` to stop it.
It goes like this:

```{r}
test_that("query works", {
  app <- presser::new_app()
  app$get("/hello", function(req, res) res$send("hello there"))
  on.exit(try(web$stop()))
  web <- presser::new_app_process(app)

  echo <- httr::content(httr::GET(web$url("/hello")))
  expect_equal(echo, "hello there")
})
```

## How can I to debug an app?

To debug an app, it is best to run it in the main R process, i.e. _not_ via `new_app_process()`.
You can add breakpoints, or `browser()` calls to your handler functions, and then invoke your app from another process.
You might find the `curl` command line tool to send HTTP requests to the app, or you can just use another R process.
Here is an example.
We will simply print the incoming request object to the screen now.
For a real debugging session you probably want to place a `browser()` command there.

```{r}
app <- presser::new_app()
app$get("/debug", function(req, res) {
  print(req)
  res$send("Got your back")
})
```

Now start the app on port 3000:

```r
app$listen(port = 3000)
```
```{r echo = FALSE}
message("Running presser web app on port 3000")
```

Connect to the app from another R or `curl` process:
```sh
curl -v http://127.0.0.1:3000/debug
```

```{r echo = FALSE}
cat('*   Trying 127.0.0.1...
* TCP_NODELAY set
* Connected to 127.0.0.1 (127.0.0.1) port 3000 (#0)
> GET /debug HTTP/1.1
> Host: 127.0.0.1:3000
> User-Agent: curl/7.54.0
> Accept: */*
>
< HTTP/1.1 200 OK
< Content-Type: text/plain
< Content-Length: 13
<
* Connection #0 to host 127.0.0.1 left intact
Got your back
')
```

Your main R session will print the incoming request:
```{r echo = FALSE}
cat('<presser_request>
method:
  get
url:
  http://127.0.0.1:3000/debug
client:
  127.0.0.1
query:
headers:
  Host: 127.0.0.1:3000
  User-Agent: curl/7.54.0
  Accept: */*
fields and methods:
  app                    # the presser_app the request belongs to
  headers                # HTTP request headers
  hostname               # server hostname, the Host header
  method                 # HTTP method of request (lowercase)
  path                   # server path
  protocol               # http or https
  query_string           # raw query string without \'?\'
  query                  # named list of query parameters
  remote_addr            # IP address of the client
  url                    # full URL of the request
  get_header(field)      # get a request header
 # see ?presser_request for details
')
```

Press `CTRL+C` or `ESC` to interrupt the app in the main session.

## Can I test asynchronous or parallel HTTP requests?

R is single threaded and a presser app runs an R interpreter, so it cannot process multiple requests at the same time.
The web server itself runs in a separate thread, and it can also process each request in a separate thread, but at any time only one request can use the R interpreter.

This is important, because sometimes test requests may take longer to process.
For example the `/delay/:secs` end point of `httpbin_app()` wait for the specified number of seconds before responding, to simulate a slow web server.
If this wait is implemented via the standard `Sys.sleep()` R function, then no other requests can be processed until the sleep is over.
To avoid this, presser can put the waiting request on hold, return from the R interpreter, and respond to other incoming requests.
Indeed, the `/delay/` end point is implemented using this feature.

However, the request thread of the web server is still busy while on hold, so to take advantage of this, you need to allow multiple threads.
The `num_threads` argument of the `$listen()` method of `presser_app` lets you specify the number of request threads the web server will use.
Similarly, the `num_threads` argument of `new_app_process()` lets you modify the number of threads.

When testing asynchronous or parallel code, that might invoke multiple, possibly delayed requests, it is best to increase the number of threads.
The code below calls the same API request concurrently, three times.
Each request takes 1 second to answer, but if the web server has more than three threads, together they'll still take about 1 second.

```{r}
web <- setup(
  presser::new_app_process(
    presser::httpbin_app(),
    opts = presser::server_opts(num_threads = 3)
  )
)
teardown(web$stop())

test_that("", {
  url <- web$url("/delay/1")
  p <- curl::new_pool()
  handles <- replicate(3, curl::new_handle(url = url))
  resps <- list()
  for (handle in handles) {
    curl::multi_add(
      handle,
      done = function(x) message("one is done"),
      fail = stop,
      pool = p
    )
  }
  st <- system.time(curl::multi_run(timeout = 5, pool = p))
  print(st)
  expect_true(st[["elapsed"]] < 3.0)
})
```
