---
title: "OAuth2.0 webfakes app"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{oauth}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(webfakes)
```

```{r child='../man/rmd-fragments/oauth2.Rmd'} 
```

In this vignette, we shall look at how to implement a flow using both apps, or the server app together with httr's OAuth tools.
In both cases we shall use the httr package as HTTP client.
For an example using the curl package instead, look at the test file `test-oauth.R` of webfakes instead.

## OAuth2.0 dance

First we need to create the resource server, and we create variables holding its different URLs.

```{r resource-server-creation}
rsapp <- local_app_process(
  oauth2_resource_app(),
  opts = server_opts(num_threads = 3)
)

rsapp

regi_url <- rsapp$url("/register")
auth_url <- rsapp$url("/authorize")
toke_url <- rsapp$url("/token")
```

Then we create the third-party app, and we create variables holding its different URLs.

```{r third-party-app-creation}
tpapp <- local_app_process(
  oauth2_third_party_app("3P app"),
  opts = server_opts(num_threads = 3)
)

tpapp

redi_url <- tpapp$url("/login/redirect")
conf_url <- tpapp$url("/login/config")
```

We then need to register the third-party app at the resource server.
In real life this is done by the admin of the third party app


```{r register}
url <- paste0(
  regi_url,
  "?name=3P%20app",
  "&redirect_uri=", redi_url
)
reg_resp <- httr::GET(url)
reg_resp
regdata <- httr::content(reg_resp)
regdata
```

Now set this data on the third-party app
In real life this is included in the config of the third party app
by its admin.

```{r register-data}
auth_data <- list(
  auth_url = auth_url,
  token_url = toke_url,
  client_id = regdata$client_id[[1]],
  client_secret = regdata$client_secret[[1]]
)

httr::POST(
  conf_url,
  body = auth_data,
  encode = "json"
)

```

##


