You can use testthat's setup files. You start the app in a setup file and also register a teardown expression for it. `local_app_process()` can do both in one go. Your `tests/testthat/setup-http.R` may look like this:

```{r eval = FALSE}
http <- webfakes::local_app_process(
  webfakes::httpbin_app(),
  .local_envir = testthat::teardown_env()
)
```

```{r include = FALSE}
http <- webfakes::new_app_process(webfakes::httpbin_app())
```

(Before testthat 3.0.0, you had to write the teardown expression in a `tests/testthat/teardown-http.R` file. That still works, but a single setup file is considered to be better practice, see [this testthat vignette](https://testthat.r-lib.org/articles/test-fixtures.html).)

In the test cases you can query the `http` app process to get the URLs you need to connect to:

```{r include = FALSE}
library(testthat)
```

```{r}
test_that("fails on 404", {
  url <- http$url("/status/404")
  response <- httr::GET(url)
  expect_error(
    httr::stop_for_status(response),
    class = "http_404"
  )
})
```

```{r, include = FALSE}
http$stop()
```

When writing your tests interactively, you may create a `http` app process in the global environment, for convenience. You can `source()` your `setup-http.R` file for this. Alternatively, you can start the app process in a helper file. See "How do I start the app when writing the tests?" just below.
