
You create a new app with `new_app()`.
This returns an object with methods to add middleware and API
endpoints to it.
For example, a simple app that returns the current time in JSON would
look like this:

```{r}
time <- presser::new_app()
time$get("/time", function(req, res) {
  res$send_json(list(time = format(Sys.time())), auto_unbox = TRUE)
})
```

Now you can start this app on a random port using `web$listen()`.
Alternatively, you can start it in a subprocess with `new_app_process()`.

```{r}
web <- presser::new_app_process(time)
web$url()
```

Use `web$url()` to query the URL of the app. For example:

```{r}
url <- web$url("/time")
httr::content(httr::GET(url))
```

`web$stop()` stops the app and the subprocess as well:

```{r}
web$stop()
web$get_state()
```

You can create your app at the beginning of your test file.
Or, if you want to use the same app in multiple test files, use a
testthat helper file.
Sometimes it useful if your users can create and use your test app,
for example to create reproducible examples.
You can include a (possibly internal) function in your package, that
creates the app.

See `?new_app()` and `?new_app_process()` for more details.
