
You create a new app with `new_app()`.
This returns an object with methods to add middleware and API
endpoints to it.
For example, a simple app that returns the current time in JSON would
look like this:

```{r}
time <- webfakes::new_app()
time$get("/time", function(req, res) {
  res$send_json(list(time = format(Sys.time())), auto_unbox = TRUE)
})
```

Now you can start this app on a random port using `web$listen()`.
Alternatively, you can start it in a subprocess with `new_app_process()`.

```{r}
web <- webfakes::new_app_process(time)
web$url()
```

Use `web$url()` to query the URL of the app. For example:

```{r}
url <- web$url("/time")
httr::content(httr::GET(url))
```

`web$stop()` stops the app and the subprocess as well:

```{r}
web$stop()
web$get_state()
```

`local_app_process()` is similar to `new_app_process()`, but it stops the server process at the end of the calling block. 
This means that the process is automatically cleaned up at the end of a `test_that()` block or at the end of the test file.

You can create your app at the beginning of your test file.
Or, if you want to use the same app in multiple test files, use a
testthat helper file.
Sometimes it useful if your users can create and use your test app,
for example to create reproducible examples.
You can include a (possibly internal) function in your package, that
creates the app.

See `?new_app()`, `?new_app_process()` and `?local_app_process` for more details.
